Declarative Integration of Interactive 3D Graphics into the World-Wide Web:\\Principles, Current Approaches, and Research Agenda \\~

### 2013-03-06 Web3D2013-Dec3D/paper.tex
new images and updates from Sandy

\TOGonlineid{XXX}
%\title{A Polyfill Approach to Declarative Integration of Interactive 3D Graphics\\ into the World-Wide Web}
\title{Declarative Integration of Interactive 3D Graphics into the World-Wide Web:\\Principles, Current Approaches, and Research Agenda}
\author{Name Lastname\thanks{e-mail:name.lastname@someemail.com}\\Some Research Institution}
\pdfauthor{Name Lastname}

\keywords{Declarative 3D, HTML, DOM, Polyfill}

@@

\begin{abstract}
With the advent of WebGL, plugin-free hardware-accelerated interactive 3D graphics has finally arrived in all major desktop Web browsers.
WebGL is an imperative solution that is tied to the functionality of rasterization API. Consequently, its usage requires a deeper understanding of the rasterization pipeline. In contrast to this stands a declarative approach with an abstract description of the 3D scene. We strongly believe that such approach is more suitable for the integration of 3D into HTML5 and related technologies such as XML (that could be used for scene construction), DOM (could be used for scene manipulation), and CSS (could be used for styling), concepts known to the millions of existing Web developers and therefore crucial for the fast adoption of 3D on the Web.

As the participants of the \textit{Declarative 3D for the Web Architecture W3C Community Group}, we explore the options for new declarative ways of incorporating 3D graphics directly into HTML to enable its use on any Web page. In this paper we present Declarative 3D principles that guide the work of the group. We describe the current state of the fundamental to our initiative experimental X3DOM and XML3D frameworks. Finally we draw the agenda for the next development stage of the Declarative 3D for the Web.
\end{abstract}

@@

Although some of these goals could also be achieved by imperative means (e.g., through the usage of WebGL \cite{Khronos-WebGL}), developments of 3D models have a long tradition of using declarative approaches, which is also in line with some of the fundamental principles of Web development. It is therefore important to explore how the experiences accumulated in two different communities, namely the Web Development and Computer Graphics communities, can be capitalized upon to achieve the long term goal of using 3D on the Web the same way as we do with 2D graphics and video today.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{images/Declarative3d.png}
  \caption{The position of the Declarative 3D approach in the current Web graphics technology ecosystem.}
  \label{fig:DeclarativeVsImperative}
  %\vspace{-10px}
\end{figure}

@@

This paper presents the current state and efforts of the Dec3D CG. In Section \ref{sec:Principles} we describe Declarative 3D principles: a number of key goals that guide the development of the DOM-based interactive 3D graphics within the Community Group.
In Section \ref{sec:Frameworks} we describe the current state of the fundamental to our initiative experimental systems: Fraunhofer's X3DOM \cite{Behr2009} and DFKI's XML3D \cite{Sons2010} declarative 3D Web publishing frameworks.
In Section \ref{sec:Agenda} we aim to shape the agenda and identify upcoming research issues for the next development stage of the Declarative 3D for the Web. Finally, in Section \ref{sec:Conclusions} we conclude our work.

\section{Declarative 3D Principles}
\label{sec:Principles}
The following goals and principles should guide the development:

@@

\begin{description}
  \item [Separation of structure from content] - Underlying the Web from its earliest days was the separation of structure from content. The concept of a paragraph specified by the $<$p$>$ tag was separate from the content in the paragraph. Declarative 3D is attempting to bring the same separation of structure from content to 3D graphics inside of web pages. The concepts such as definitions of 3D objects, transformations, materials, etc. should be implemented in a declarative 3D description as an extension to HTML5 using any existing or future extension mechanism.
  \item [Separation of content from style] - One of the principles of the current Web is also the separation of content and style, most notably through CSS. The successful integration of SVG with HTML was made much easier due to the fact that SVG was already following this principle.
      The aim of the CG is to extend the use of CSS for styling 3D graphics. One example can be the use of the latest CSS3 3D Transforms to allow to manipulate not only 2D but also 3D objects in 3D space.
  \item [Use of the Document Object Model] - The Document Object Model (DOM) is a platform- and language-neutral interface that allows programs and scripts to dynamically access and update the content, structure and style of Web documents \cite{W3C-DOM}. Declarative 3D should use DOM API to examine and modify elements on the 3D scene and their attributes by simply reading and setting their properties.
      As DOM provides access to user actions (e.g., pressing a key or clicking a mouse button), it should also be used as a main interface to interact with 3D content.
\end{description}

@@

\newpage

\section{Declarative 3D Experimental Frameworks}
\label{sec:Frameworks}
As we have already mentioned, our Community Group has been formed to determine the requirements, options, and use cases for the declarative integration of interactive 3D graphics capabilities into the Web technology stack which will provide a foundation for future standardization.
While the standardization is our goal, we still need platforms allowing for the experimentation and for the assessment of our design decisions. We also need to reach out to Web developers who could provide us with valuable feedback as early as possible.

@@

\subsection{X3DOM \& XML3D}
Introduction to this section: Fraunhofer's X3DOM \cite{Behr2009} is an open-source framework that allows including X3D \cite{Web3D-X3D} elements as part of any HTML DOM tree... DFKI's XML3D \cite{Sons2010} is a similar approach...

@@

\newpage

\subsection{Polyfill Approach}
Recently, the CG proposed to measure a level of integration for 3D graphics in terms of W3C technologies (DOM, CSS etc) \cite{Dec3D-LevelsOfIntegration}. The proposal aims to explain existing and possible integration levels and finally to identify an integration level the CG is aiming for.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{images/Dec3D-Architecture.png}
  \caption{Declarative 3D polyfill runtime architecture.}
\end{figure}

\begin{figure*}[ht]
  \centering
  \includegraphics[width=\linewidth]{images/Demos.png}
  \caption{ToDo: Some latest examples of X3DOM and XML3D based Web applications.}
  \vspace{-7px}
\end{figure*}

@@

\newpage

\section{Declarative 3D Agenda}
\label{sec:Agenda}
During the 1st International Workshop on Declarative 3D for the Web Architecture (Dec3D2012) \cite{Dec3D2012} as well as during more informal group meetings at Web3D in 2011, 2012 and SIGGRAPH in 2011, 2012, the members and supporters of the Declarative 3D W3C Community Group agreed that what is needed to make the effort successful and for the W3C to adopt/develop a declarative 3D standard is:

@@

\paragraph{Outreach \& Killer App}
The CG needs to continue its outreach activities through the high quality demonstrations of the Declarative 3D philosophy using the experimental X3DOM and XML3D frameworks. It should also consider focusing on one killer application of Declarative 3D, an application so necessary or desirable that it proves the core value of the Declarative approach and that could substantially increase the visibility of the CG's efforts.

@@

\newpage

\section{Conclusions}
\label{sec:Conclusions}
While WebGL 3D imperative graphics API in the Web context is getting more and more traction, we are still missing an easy way to add interactive high-level declarative 3D objects to the HTML to allow anyone to easily create, share, and experience interactive 3D graphics – with possibly wide ranging effects similar to those caused by the broad availability of video on the Web.

>>>>

\TOGonlineid{XXX}

%\title{A Polyfill Approach to Declarative Integration of Interactive 3D Graphics\\ into the World-Wide Web}
\title{Declarative Integration of Interactive 3D Graphics into the World-Wide Web:\\Principles, Current Approaches, and Research Agenda
\\~} %some space that we don't get over 4 pages in the final version
\author{Name Lastname\thanks{e-mail:name.lastname@someemail.com}\\Some Research Institution}
\pdfauthor{Name Lastname}

\keywords{Declarative 3D, HTML5, DOM Integration, Polyfill}

@@

\begin{abstract}
With the advent of WebGL, plugin-free hardware-accelerated interactive 3D graphics has finally arrived in all major desktop Web browsers.
WebGL is an imperative solution that is tied to the functionality of rasterization API. Consequently, its usage requires a deeper understanding of the rasterization pipeline. In contrast to this stands a declarative approach with an abstract description of the 3D scene. We strongly believe that such approach is more suitable for the integration of 3D into HTML5 and related technologies such as XML (that could be used for scene construction), DOM (could be used for scene manipulation), and CSS (could be used for styling), concepts known to the millions of existing Web developers and therefore crucial for the fast adoption of 3D on the Web.\\
As the participants of the \textit{Declarative 3D for the Web Architecture W3C Community Group}, we explore the options for new declarative ways of incorporating 3D graphics directly into HTML to enable its use on any Web page. In this paper we present Declarative 3D principles that guide the work of the group. We describe the current state of the fundamental to our initiative experimental X3DOM and XML3D frameworks. Finally we draw the agenda for the next development stage of the Declarative 3D for the Web.
\end{abstract}

@@

Although some of these goals could also be achieved by imperative means (e.g., through the usage of WebGL \cite{WebGL12}), developments of 3D models have a long tradition of using declarative approaches, which is also in line with some of the fundamental principles of Web development. It is therefore important to explore how the experiences accumulated in two different communities, namely the Web Development and Computer Graphics communities, can be capitalized upon to achieve the long term goal of using 3D on the Web the same way as we do with 2D graphics and video today.

\begin{figure}%[ht]
  \centering
  \includegraphics[width=\linewidth]{images/Declarative3d.png}
  \caption{The position of the Declarative 3D approach in the current Web graphics technology ecosystem.}
  \label{fig:DeclarativeVsImperative}
  %\vspace{-10px}
\end{figure}

@@

This paper presents the current state and efforts of the Dec3D CG. In Section \ref{sec:Principles} we describe Declarative 3D principles: a number of key goals that guide the development of the DOM-based interactive 3D graphics within the Community Group.
In Section \ref{sec:Frameworks} we describe the current state of the fundamental to our initiative experimental systems: Fraunhofer IGD's X3DOM \cite{Behr2009} and DFKI's XML3D \cite{Sons2010} declarative 3D Web publishing frameworks.
In Section \ref{sec:Agenda} we aim to shape the agenda and identify upcoming research issues for the next development stage of the Declarative 3D for the Web. Finally, in Section \ref{sec:Conclusions} we conclude our work along with directions for future work.

\section{Declarative 3D Principles}
\label{sec:Principles}

The following goals and principles should guide the development:

@@

\begin{description}
  \item [Separation of structure from content] - Underlying the Web from its earliest days was the separation of structure from content. The concept of a paragraph specified by the \verb|<p>| tag was separate from the content in the paragraph. Declarative 3D is attempting to bring the same separation of structure from content to 3D graphics inside of web pages. The concepts such as definitions of 3D objects, transformations, materials, etc. should be implemented in a declarative 3D description as an extension to HTML5 using any existing or future extension mechanism.
  \item [Separation of content from style] - One of the principles of the current Web is also the separation of content and style, most notably through CSS. The successful integration of SVG with HTML was made much easier due to the fact that SVG was already following this principle.
      The aim of the CG is to extend the use of CSS for styling 3D graphics. One example can be the use of the latest CSS3 3D Transforms to allow to manipulate not only 2D but also 3D objects in 3D space.
  \item [Use of the Document Object Model] - The Document Object Model (DOM) is a platform- and language-neutral interface that allows programs and scripts to dynamically access and update the content, structure and style of Web documents \cite{W3C-DOM}. Declarative 3D should use DOM API to examine and modify elements on the 3D scene and their attributes by simply reading and setting their properties.
      As DOM provides access to user actions (e.g., pressing a key or clicking a mouse button), it should also be used as a main interface to interact with 3D content.
\end{description}

@@

\paragraph{Leveraging Web Development Infrastructure}
The Web has become the new application development platform for our time. Countless rich internet applications are being created at a dizzying pace. The ``apps'' world for mobile applications has also become a huge phenomenon. The single most important technology, among many, that has emerged to enable this explosion of creative applications is JavaScript. JavaScript has evolved from a toy language into a robust and richly nuanced tool that is the basis for much of the recent explosion in Web applications.
Declarative 3D, by living within the structure of Web applications is poised to leverage the huge collection of tools and infrastructure that already exists like e.g. the JavaScript library jQuery \cite{jquery}. The ability to take existing Web debuggers, editors, and viewers and use these tools to create, edit and debug 3D content is a tremendous benefit and just begins to scratch the surface of development tools, created for Web application development, but which we can use for 3D content development.

\section{Declarative 3D Experimental Frameworks}
\label{sec:Frameworks}

As we have already mentioned, our Community Group has been formed to determine the requirements, options, and use cases for the declarative integration of interactive 3D graphics capabilities into the Web technology stack which will provide a foundation for future standardization.
While the standardization is our goal, we still need platforms allowing for the experimentation and for the assessment of our design decisions. We also need to reach out to Web developers who could provide us with valuable feedback as early as possible.

@@

\subsection{X3DOM \& XML3D}
Introduction to this section: Fraunhofer's X3DOM \cite{Behr2009,Behr12} is an open-source framework that allows including X3D \cite{Web3D-X3D} elements as part of any HTML DOM tree... DFKI's XML3D \cite{Sons2010} is a similar approach...

@@

% TODO
Efficient 3D transmission formats XXX \cite{Behr12}, and are also a major topic at the Khronos Group \cite{Trevett12}.

\subsection{Polyfill Approach}

\begin{figure}%[ht]
  \centering
  \includegraphics[width=\linewidth]{images/Dec3D-Architecture.png}
  \caption{Declarative 3D polyfill runtime architecture.}
\end{figure}

Recently, the CG proposed to measure a level of integration for 3D graphics in terms of W3C technologies (DOM, CSS etc) \cite{Dec3D-LevelsOfIntegration}. The proposal aims to explain existing and possible integration levels and finally to identify an integration level the CG is aiming for.

@@

@@

\begin{figure*}%[ht]
  \centering
  \includegraphics[width=\linewidth]{images/x3dom_apps.png}
  \caption{Some X3DOM applications. From left to right: walktrough of Roman heritage site; cultural heritage object explorer with metadata; web-based design review application with annotation markers; rendering big data sets, here the Boeing 777, in the Web browser using an ``out-of-browser'' approach. (The Boeing 777 commercial airplane model is courtesy of Boeing. License granted by David Kasik, Boeing.)}
  \vspace{-0.2cm}
\end{figure*}

\section{Declarative 3D Agenda}

\label{sec:Agenda}
During the 1st International Workshop on Declarative 3D for the Web Architecture (Dec3D2012) \cite{Dec3D2012} as well as during more informal group meetings at Web3D in 2011, 2012 and SIGGRAPH in 2011, 2012, the members and supporters of the Declarative 3D W3C Community Group agreed that what is needed to make the effort successful and for the W3C to adopt/develop a declarative 3D standard is:

@@

\paragraph{Outreach \& Exemplar Applications}
The CG needs to continue its outreach activities through the high quality demonstrations of the Declarative 3D philosophy using the experimental X3DOM and XML3D frameworks. %It should also consider focusing on one killer application of Declarative 3D, an application so necessary or desirable that it proves the core value of the Declarative approach and that could substantially increase the visibility of the CG's efforts.
In this regard, the CG should identify several applications, each requiring and demonstrating different capabilities of Declarative 3D. For example, one application could require huge 3D datasets which are impractical for inclusion directly into the DOM; another could require real-time control of an external system (inter process communication); another could integrate with a complex data base; another an illustration of complex data (information visualization) without a physical analog and another a more traditional scientific visualization. Key to selection of these applications is the use and demonstration of a Declarative 3D requirement.

@@

\section{Conclusions}

\label{sec:Conclusions}
While WebGL 3D imperative graphics API in the Web context is getting more and more traction, we are still missing an easy way to add interactive high-level declarative 3D objects to the HTML to allow anyone to easily create, share, and experience interactive 3D graphics – with possibly wide ranging effects similar to those caused by the broad availability of video on the Web.
